---
output:
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# SpThreshold: Replication Threshold for Bayesian Multilevel Models of Areal Data

## SpThreshold_Example

[1] Initialize package and set reproducibility seed:
```{r}
library(SpThreshold)
set.seed(2026)
```

[2] Generate a random connected adjacency matrix:
```{r}
n<-100  #Number of spatial units
W<-create_random_W(n_locs = n)
```

[3] Set covariance parameters:

* Within-area residual variance, spatial random-effect variance, spatial correlation:
```{r}
sigma2<-0.50
tau2<-0.50
rho<-0.70
```

[4] Compute $m^*$ under three covariate structures:

* C1: covariate varies only within areas ($\text{Var}(\bar{\text{x}}_{i \cdot}) \approx 0$):
```{r}
m_obs<-20  #Observations per area
N<-n*m_obs
loc<-rep(1:n,
         each = m_obs)
X_c1<-rnorm(n = N)
X_c1<-(X_c1 - mean(X_c1))/sqrt(mean((X_c1 - mean(X_c1))^2))  #Standardize
res_c1<-m_star(W = W,
               sigma2 = sigma2,
               tau2 = tau2,
               rho = rho,
               X = X_c1,
               loc = loc)
res_c1$m_star
```

* C2: covariate varies both within and between areas:
```{r}
loc_means<-rnorm(n = n)
X_c2<-rnorm(n = N,
            mean = loc_means[loc],
            sd = 1.00)
X_c2<-(X_c2 - mean(X_c2))/sqrt(mean((X_c2 - mean(X_c2))^2))
res_c2<-m_star(W = W,
               sigma2 = sigma2,
               tau2 = tau2,
               rho = rho,
               X = X_c2,
               loc = loc)
res_c2$m_star
```

* C3: covariate varies only between areas ($\text{Var}(\text{x}_{ij} \mid i) = 0$, denominator vanishes, bound is infinite):
```{r}
X_c3<-loc_means[loc]
X_c3<-(X_c3 - mean(X_c3))/sqrt(mean((X_c3 - mean(X_c3))^2))
res_c3<-m_star(W = W,
               sigma2 = sigma2,
               tau2 = tau2,
               rho = rho,
               X = X_c3,
               loc = loc)
res_c3$m_star
```

[5] Vary $\rho$ under C2:
```{r}
rho_seq<-c(0.05, 0.25, 0.50, 0.75, 0.95)
m_star_rho<-sapply(rho_seq,
                   function(r) m_star(W = W,
                                      sigma2 = 0.50,
                                      tau2 = 0.50,
                                      rho = r,
                                      X = X_c2,
                                      loc = loc)$m_star)
data.frame(rho = rho_seq,
           m_star = m_star_rho)
```

[6] Vary $\kappa = \tau^2 / \sigma^2$ under C2:
```{r}
k_seq<-c(0.05/0.95, 0.50/0.50, 0.95/0.05)
m_star_k<-sapply(k_seq,
                 function(k){
                    tau2_k<-k/(1.00 + k)
                    sigma2_k<-1.00/(1.00 + k)
                    m_star(W = W,
                           sigma2 = sigma2_k,
                           tau2 = tau2_k,
                           rho = 0.70,
                           X = X_c2,
                           loc = loc)$m_star
                    })
data.frame(k = round(k_seq, 3),
           m_star = m_star_k)
```

[7] Construct $W$ from a shapefile:
```{r eval=FALSE}
library(sf)
nc<-st_read(system.file("shape/nc.shp",
                        package = "sf"))
W_nc<-create_W_from_shapefile(polygons = nc)
```
