---
output:
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# SpThreshold: Replication Threshold for Bayesian Multilevel Models of Areal Data

## Part I: Replication threshold via m_star()

[1] Initialize package and set reproducibility seed:
```{r}
library(SpThreshold)
set.seed(2026)
```

[2] Generate a random connected adjacency matrix:
```{r}
n<-100  #Number of spatial units
W<-create_random_W(n_locs = n)
```

[3] Set covariance parameters:

* Within-area residual variance, spatial random-effect variance, spatial correlation:
```{r}
sigma2<-0.50
tau2<-0.50
rho<-0.70
```

[4] Compute $m^*$ under three covariate structures:

* C1: covariate varies only within areas ($\text{Var}(\bar{\text{x}}_{i \cdot}) \approx 0$):
```{r}
m_obs<-20  #Observations per area
N<-n*m_obs
loc<-rep(1:n,
         each = m_obs)
X_c1<-rnorm(n = N)
X_c1<-(X_c1 - mean(X_c1))/sqrt(mean((X_c1 - mean(X_c1))^2))  #Standardize
res_c1<-m_star(W = W,
               sigma2 = sigma2,
               tau2 = tau2,
               rho = rho,
               X = X_c1,
               loc = loc)
res_c1$m_star
```

* C2: covariate varies both within and between areas:
```{r}
loc_means<-rnorm(n = n)
X_c2<-rnorm(n = N,
            mean = loc_means[loc],
            sd = 1.00)
X_c2<-(X_c2 - mean(X_c2))/sqrt(mean((X_c2 - mean(X_c2))^2))
res_c2<-m_star(W = W,
               sigma2 = sigma2,
               tau2 = tau2,
               rho = rho,
               X = X_c2,
               loc = loc)
res_c2$m_star
```

* C3: covariate varies only between areas ($\text{Var}(\text{x}_{ij} \mid i) = 0$, denominator vanishes, bound is infinite):
```{r}
X_c3<-loc_means[loc]
X_c3<-(X_c3 - mean(X_c3))/sqrt(mean((X_c3 - mean(X_c3))^2))
res_c3<-m_star(W = W,
               sigma2 = sigma2,
               tau2 = tau2,
               rho = rho,
               X = X_c3,
               loc = loc)
res_c3$m_star
```

[5] Vary $\rho$ under C2:
```{r}
rho_seq<-c(0.05, 0.25, 0.50, 0.75, 0.95)
m_star_rho<-sapply(rho_seq,
                   function(r) m_star(W = W,
                                      sigma2 = 0.50,
                                      tau2 = 0.50,
                                      rho = r,
                                      X = X_c2,
                                      loc = loc)$m_star)
data.frame(rho = rho_seq,
           m_star = m_star_rho)
```

[6] Vary $\kappa = \tau^2 / \sigma^2$ under C2:
```{r}
k_seq<-c(0.05/0.95, 0.50/0.50, 0.95/0.05)
m_star_k<-sapply(k_seq,
                 function(k){
                    tau2_k<-k/(1.00 + k)
                    sigma2_k<-1.00/(1.00 + k)
                    m_star(W = W,
                           sigma2 = sigma2_k,
                           tau2 = tau2_k,
                           rho = 0.70,
                           X = X_c2,
                           loc = loc)$m_star
                    })
data.frame(k = round(k_seq, 3),
           m_star = m_star_k)
```

[7] Construct $W$ from a shapefile:

```{r}
library(sf)

# please replace with the path to your own shapefile; 
# here we use the example shapefile bundled with sf
shp_path <- system.file("shape/nc.shp", package = "sf")
polygons <- st_read(shp_path, quiet = TRUE)

W_nc <- create_W_from_shapefile(polygons = polygons)
```

[8] Vary the tolerance $\gamma$ under C2:
```{r}
gamma_seq<-c(0.01, 0.05, 0.10, 0.20)
m_star_gamma<-sapply(gamma_seq,
                     function(g) m_star(W = W,
                                        sigma2 = 0.50,
                                        tau2 = 0.50,
                                        rho = 0.70,
                                        X = X_c2,
                                        loc = loc,
                                        gamma = g)$m_star)
data.frame(gamma = gamma_seq,
           m_star = m_star_gamma)
```


## Part II: Model fitting via spfit()

`spfit()` fits the multilevel model from the manuscript via MCMC, with either a Leroux CAR prior (`model_indicator = 1`) or an iid prior (`model_indicator = 0`) on the area-level random effects. The C++ implementation has been verified to target the same posterior as the R reference to within Monte Carlo error.

[1] Generate a small synthetic dataset. We use a C2-style covariate (within- and between-area variation) here for better illustration.
```{r}
set.seed(2026)
n_locs<-50
m_per<-10
N<-n_locs*m_per

W_small<-as.matrix(create_random_W(n_locs))
loc_small<-rep(1:n_locs, each = m_per)

beta_true<-c(0.50, 1.50)
sigma2_true<-0.50
tau2_true<-0.50
rho_true<-0.50

D<-diag(rowSums(W_small))
Q_true<-rho_true*(D - W_small) + (1.00 - rho_true)*diag(n_locs)
theta_true<-as.numeric(t(chol(solve(Q_true)*tau2_true)) %*% rnorm(n_locs))
theta_true<-theta_true - mean(theta_true)

loc_means_x<-rnorm(n_locs)
x_small<-rnorm(N, mean = loc_means_x[loc_small], sd = 1.00)
x_small<-(x_small - mean(x_small))/sqrt(mean((x_small - mean(x_small))^2))
X_small<-cbind(1, x_small)
y_small<-as.numeric(X_small %*% beta_true + theta_true[loc_small]) +
         rnorm(N, 0, sqrt(sigma2_true))
```

[2] Fit the spatial model:
```{r}
fit_sp<-spfit(y = y_small,
              X = X_small,
              loc = loc_small,
              W = W_small,
              model_indicator = 1,
              mcmc_samples = 15000,
              burnin = 5000)
```

[3] Discard burn-in and summarize the posterior:
```{r}
keep<-5001:15000
data.frame(parameter = c("beta0", "beta1", "sigma2", "tau2", "rho"),
           truth = c(beta_true, sigma2_true, tau2_true, rho_true),
           posterior_mean = c(mean(fit_sp$beta[1, keep]),
                              mean(fit_sp$beta[2, keep]),
                              mean(fit_sp$sigma2[keep]),
                              mean(fit_sp$tau2[keep]),
                              mean(fit_sp$rho[keep])),
           posterior_sd = c(sd(fit_sp$beta[1, keep]),
                            sd(fit_sp$beta[2, keep]),
                            sd(fit_sp$sigma2[keep]),
                            sd(fit_sp$tau2[keep]),
                            sd(fit_sp$rho[keep])))
```

[4] Fit the nonspatial model on the same data:
```{r}
fit_ns<-spfit(y = y_small,
              X = X_small,
              loc = loc_small,
              W = W_small,
              model_indicator = 0,
              mcmc_samples = 15000,
              burnin = 5000)
```

[5a] Compare $\beta_1$ inference between the two models:
```{r}
data.frame(model = c("spatial", "nonspatial"),
           posterior_mean = c(mean(fit_sp$beta[2, keep]),
                              mean(fit_ns$beta[2, keep])),
           posterior_sd = c(sd(fit_sp$beta[2, keep]),
                            sd(fit_ns$beta[2, keep])),
           q025 = c(quantile(fit_sp$beta[2, keep], 0.025, names = FALSE),
                    quantile(fit_ns$beta[2, keep], 0.025, names = FALSE)),
           q975 = c(quantile(fit_sp$beta[2, keep], 0.975, names = FALSE),
                    quantile(fit_ns$beta[2, keep], 0.975, names = FALSE)))
```

[5b] Sanity check: by definition, $m^*$ is the smallest $m$ where the relative difference in posterior variance of $\beta_1$ falls below the tolerance $\gamma$. Compute that relative difference here and confirm it's below $\gamma = 0.05$:
```{r}
var_sp<-var(fit_sp$beta[2, keep])
var_ns<-var(fit_ns$beta[2, keep])
rel_diff<-abs(var_sp - var_ns)/var_ns

data.frame(var_sp = var_sp,
           var_ns = var_ns,
           rel_diff = rel_diff,
           gamma = 0.05,
           within_threshold = rel_diff < 0.05)
```

[6] Connect the two parts: compute $m^*$ for this C2 covariate and check whether the available replication exceeds it. When $m \geq m^*$, the spatial and nonspatial posteriors for $\beta_1$ should be effectively equivalent.
```{r}
xbar_small<-as.numeric(tapply(x_small, loc_small, mean))
res<-m_star(W = W_small,
            sigma2 = sigma2_true,
            tau2 = tau2_true,
            rho = rho_true,
            X = x_small,
            loc = loc_small)
data.frame(m_star = res$m_star,
           m_actual = m_per,
           sufficient = m_per >= res$m_star)
```

[7] The data frame interface is also supported and produces identical output. Below uses the same data via a data frame:
```{r}
df_small<-data.frame(y = y_small, x = x_small, area = loc_small)

set.seed(7)
fit_df<-spfit(data = df_small,
              formula = y ~ x,
              area = "area",
              W = W_small,
              model_indicator = 1,
              mcmc_samples = 15000,
              burnin = 5000)

set.seed(7)
fit_vec<-spfit(y = y_small,
               X = X_small,
               loc = loc_small,
               W = W_small,
               model_indicator = 1,
               mcmc_samples = 15000,
               burnin = 5000)

identical(fit_df$beta, fit_vec$beta)
```


